### 2023-02-07 draft.tex
Deploying to gh-pages from @ PoisotLab/ms_interpreting_probabilities@5a85138104cfa2abd413f46f7fec3106326b9c3d 🚀

Cataloging species interactions across space is a gargantuan task. At
the core of this challenge lies the spatiotemporal variability of
ecological networks (Poisot \emph{et al.} 2012, 2015), which makes
documenting the location and timing of interactions difficult. Indeed,
it is not sufficient to know if two species have the biological capacity
to interact to infer the realization of their interaction at a specific
time and space {[}TK{]}. Taking food webs as an example, a predator and
its potential prey must first co-occur on the same territory in order
for a trophic interaction to take place (Blanchet \emph{et al.} 2020).
They must then encounter, which is conditional on their relative
abundances in the ecosystem and the matching of their phenology (Poisot
\emph{et al.} 2015). Finally, the interaction occurs only if the
predators have a desire to consume their prey and are able to capture
and ingest them (Pulliam 1974). Environmental (e.g.~temperature and
presence of shelters) and biological (e.g.~physiological state of both
species and availability of other prey species) factors contribute to
this variability by impacting species co-occurrence {[}TK{]} and the
realization of their interactions {[}TK{]}. In this context, it is
unsurprising that computational methods are being developed in ecology
to help alleviate the colossal sampling efforts required to document
species interactions across time and space (Strydom \emph{et al.} 2021).
Having a better portrait of species interactions and the emerging
structure of food webs is important since it lays the groundwork for
understanding the functioning {[}TK{]}, dynamics {[}TK{]}, and
resilience {[}TK{]} of ecosystems worldwide.

The recognition of the variability of species interactions and the
emergence of numerical methods have led ecologists to rethink their
representation of ecological networks, slowly moving from a binary to a
probabilistic view of species interactions (Poisot \emph{et al.} 2016).
This has several benefits. For example, probabilities represent the
limit of our knowledge about species interactions and can indicate the
expected frequency of two species interacting with each other {[}TK{]}.
They are also very helpful in predictive models when modeling the
spatial distribution of species (Cazelles \emph{et al.} 2016) and the
temporal variability of ecological networks {[}TK{]}, generating new
ecological data (e.g., Strydom \emph{et al.} 2022), and identifying
priority sampling locations (see Andrade-Pacheco \emph{et al.} 2020 for
an ecological example of a sampling optimization problem). Moreover, the
high rate of false negatives in ecological network data, resulting from
the difficulty of witnessing interactions between rare species, makes it
hard to interpret non-observations of species interactions ecologically
(Catchen \emph{et al.} 2023). Using probabilities instead of yes-no
interactions accounts for these observation errors; in that case, only
forbidden interactions (Jordano \emph{et al.} 2003; Olesen \emph{et al.}
2010) would have a probability value of zero (but see Gonzalez-Varo \&
Traveset 2016).

@@

@@

i = predator j = prey C = co-occurrence A = area N = relative abundance
E = environment (including network) t = time

@@

\hypertarget{refs}{}
\begin{CSLReferences}{1}{0}
\leavevmode\hypertarget{ref-Andrade-Pacheco2020FinHot}{}%
Andrade-Pacheco, R., Rerolle, F., Lemoine, J., Hernandez, L., Meïté, A.,
Juziwelo, L., \emph{et al.} (2020). Finding hotspots: Development of an
adaptive spatial sampling approach. \emph{Scientific Reports}, 10,
10939.

\leavevmode\hypertarget{ref-Blanchet2020CooNot}{}%
Blanchet, F.G., Cazelles, K. \& Gravel, D. (2020). Co-occurrence is not
evidence of ecological interactions. \emph{Ecology Letters}, 23,
1050--1063.

\leavevmode\hypertarget{ref-Catchen2023MisLin}{}%
Catchen, M.D., Poisot, T., Pollock, L.J. \& Gonzalez, A. (2023). The
missing link: Discerning true from false negatives when sampling species
interaction networks.

\leavevmode\hypertarget{ref-Cazelles2016TheSpe}{}%
Cazelles, K., Araujo, M.B., Mouquet, N. \& Gravel, D. (2016). A theory
for species co-occurrence in interaction networks. \emph{Theoretical
Ecology}, 9, 39--48.

\leavevmode\hypertarget{ref-Gonzalez-Varo2016LabLim}{}%
Gonzalez-Varo, J.P. \& Traveset, A. (2016). The Labile Limits of
Forbidden Interactions. \emph{Trends in Ecology \& Evolution}, 31,
700--710.

\leavevmode\hypertarget{ref-Jordano2003InvPro}{}%
Jordano, P., Bascompte, J. \& Olesen, J.M. (2003). Invariant properties
in coevolutionary networks of plantanimal interactions. \emph{Ecology
Letters}, 6, 69--81.

@@

\leavevmode\hypertarget{ref-Olesen2010MisFor}{}%
Olesen, J.M., Bascompte, J., Dupont, Y.L., Elberling, H., Rasmussen, C.
\& Jordano, P. (2010). Missing and forbidden links in mutualistic
networks. \emph{Proceedings of the Royal Society B: Biological
Sciences}, 278, 725--732.

\leavevmode\hypertarget{ref-Poisot2012DisSpe}{}%
Poisot, T., Canard, E., Mouillot, D., Mouquet, N. \& Gravel, D. (2012).
The dissimilarity of species interaction networks. \emph{Ecology
Letters}, 15, 1353--1361.

@@

\leavevmode\hypertarget{ref-Poisot2015SpeWhy}{}%
Poisot, T., Stouffer, D.B. \& Gravel, D. (2015). Beyond species: Why
ecological interaction networks vary through space and time.
\emph{Oikos}, 124, 243--251.

\leavevmode\hypertarget{ref-Pulliam1974TheOpt}{}%
Pulliam, H.R. (1974). On the Theory of Optimal Diets. \emph{The American
Naturalist}, 108, 59--74.

@@

\leavevmode\hypertarget{ref-Strydom2021RoaPre}{}%
Strydom, T., Catchen, M.D., Banville, F., Caron, D., Dansereau, G.,
Desjardins-Proulx, P., \emph{et al.} (2021). A roadmap towards
predicting species interaction networks (across space and time).
\emph{Philosophical Transactions of the Royal Society B-Biological
Sciences}, 376, 20210063.

>>>>

Cataloging species interactions across space is a gargantuan task. At
the core of this challenge lies the spatiotemporal variability of
ecological networks (\textbf{Poisot2012DisSpe?};
\textbf{Poisot2015SpeWhy?}), which makes documenting the location and
timing of interactions difficult. Indeed, it is not sufficient to know
that two species have the biological capacity to interact to infer the
realization of their interaction at a specific time and space (Dunne
2006). Taking food webs as an example, a predator species and its
potential prey must first co-occur on the same territory in order for a
trophic interaction to take place (\textbf{Blanchet2020CooNot?}). They
must then encounter, which is conditional on their relative abundances
in the ecosystem and the matching of their phenology
(\textbf{Poisot2015SpeWhy?}). Finally, the interaction occurs only if
the predators have a desire to consume their prey and are able to
capture and ingest them (\textbf{Pulliam1974TheOpt?}). Environmental
(e.g.~temperature and presence of shelters) and biological
(e.g.~physiological state of both species and availability of other prey
species) factors contribute to this variability by impacting species
co-occurrence (Araujo \emph{et al.} 2011) and the realization of their
interactions (\textbf{Poisot2015SpeWhy?}). In this context, it is
unsurprising that computational methods are being developed in ecology
to help alleviate the colossal sampling efforts required to document
species interactions across time and space
(\textbf{Strydom2021RoaPre?}). Having a better portrait of species
interactions and the emerging structure of their food webs is important
since it lays the groundwork for understanding the functioning,
dynamics, and resilience of ecosystems worldwide (e.g., Proulx \emph{et
al.} 2005; Pascual \emph{et al.} 2006; Delmas \emph{et al.} 2019).

The recognition of the intrinsic variability of species interactions and
the emergence of numerical methods have led ecologists to rethink their
representation of ecological networks, slowly moving from a binary to a
probabilistic view of species interactions (Poisot \emph{et al.} 2016).
This has several benefits. For example, probabilities represent the
limit of our knowledge about species interactions and can inform us
about the expected number of interactions and emerging network
properties despite this limited knowledge (Poisot \emph{et al.} 2016).
They are also very helpful in predicting the spatial distribution of
species within networks (\textbf{Cazelles2016TheSpe?}) and the temporal
variability of interactions (Poisot \emph{et al.} 2015), generating new
ecological data (e.g., Strydom \emph{et al.} 2022), and identifying
priority sampling locations of species interactions (see
\textbf{Andrade-Pacheco2020FinHot?} for an ecological example of a
sampling optimization problem). Moreover, the high rate of false
negatives in ecological network data, resulting from the difficulty of
witnessing interactions between rare species, makes it hard to interpret
non-observations of species interactions ecologically
(\textbf{Catchen2023MisLin?}). Using probabilities instead of yes-no
interactions accounts for these observation errors; in that case, only
forbidden interactions (\textbf{Jordano2003InvPro?};
\textbf{Olesen2010MisFor?}) would have a probability value of zero (but
see \textbf{Gonzalez-Varo2016LabLim?}). Many measures have been
developed to describe the structure (Poisot \emph{et al.} 2016) and
diversity (Ohlmann \emph{et al.} 2019; Godsoe \emph{et al.} n.d.) of
probabilistic interactions, which shows the potential of this framework
in the study of a variety of ecological phenomena.

@@

Papers: Guimarães (2020)

@@

i = predator j = prey C = co-occurrence (and relative abundance?) A =
area N = relative abundance (or network?) E = environment (including
network?) t = time

@@

\hypertarget{refs}{}
\begin{CSLReferences}{1}{0}
\leavevmode\hypertarget{ref-Araujo2011UsiSpe}{}%
Araujo, M.B., Rozenfeld, A., Rahbek, C. \& Marquet, P.A. (2011). Using
species co-occurrence networks to assess the impacts of climate change.
\emph{Ecography}, 34, 897--908.

\leavevmode\hypertarget{ref-Delmas2019AnaEco}{}%
Delmas, E., Besson, M., Brice, M.-H., Burkle, L.A., Dalla Riva, G.V.,
Fortin, M.-J., \emph{et al.} (2019). Analysing ecological networks of
species interactions. \emph{Biological Reviews}, 94, 16--36.

\leavevmode\hypertarget{ref-Dunne2006Network}{}%
Dunne, J.A. (2006). The Network Structure of Food Webs. In:
\emph{Ecological networks: Linking structure and dynamics} (eds. Dunne,
J.A. \& Pascual, M.). Oxford University Press, pp. 27--86.

\leavevmode\hypertarget{ref-GodsoeSpeInt}{}%
Godsoe, W., Murray, R. \& Iritani, R. (n.d.). Species interactions and
diversity: A unified framework using Hill numbers. \emph{Oikos}, n/a,
e09282.

\leavevmode\hypertarget{ref-Guimaraes2020StrEco}{}%
Guimarães, P.R. (2020). The Structure of Ecological Networks Across
Levels of Organization. \emph{Annual Review of Ecology, Evolution, and
Systematics}, 51, 433--460.

@@

\leavevmode\hypertarget{ref-Ohlmann2019DivInd}{}%
Ohlmann, M., Miele, V., Dray, S., Chalmandrier, L., O'Connor, L. \&
Thuiller, W. (2019). Diversity indices for ecological networks: A
unifying framework using Hill numbers. \emph{Ecology Letters}, 22,
737--747.

\leavevmode\hypertarget{ref-Pascual2006EcoNet}{}%
Pascual, M., Dunne, J.A. \& Dunne, J.A. (2006). \emph{Ecological
Networks: Linking Structure to Dynamics in Food Webs}. Oxford University
Press, USA.

@@

\leavevmode\hypertarget{ref-Poisot2015Speciesa}{}%
Poisot, T., Stouffer, D.B. \& Gravel, D. (2015). Beyond species: Why
ecological interaction networks vary through space and time.
\emph{Oikos}, 124, 243--251.

\leavevmode\hypertarget{ref-Proulx2005NetThi}{}%
Proulx, S.R., Promislow, D.E.L. \& Phillips, P.C. (2005). Network
thinking in ecology and evolution. \emph{Trends in Ecology \&
Evolution}, SPECIAL issue: BUMPER book REVIEW, 20, 345--353.

@@
